---
title: "Capstone project ~ Data Wrangling"
author: "Elena Tortosa"
date: "2/1/2019"
output: pdf_document
---

## Libraries and work directory

I loaded all required libraries and set the work directory.

```{r message=FALSE}
library(tidyr)
library(dplyr)
setwd("/Users/Tortosae/Desktop/Data science course/Capstone_project")
```

## Data tables

All data was provided in 14 different excel sheets. I saved all as .csv and loaded them in R. 
They were named with a number (df#) followed by the name of the perturbation + treatment each table contains.

```{r message=FALSE}
df1_cd3cd28 <- read.table(file="1.cd3cd28.csv",sep=",", header=TRUE)
df2_cd3cd28icam2 <- read.table(file="2. cd3cd28icam2.csv",sep=",", header=TRUE)
df3_cd3cd28aktinhib <- read.table(file="3. cd3cd28aktinhib.csv",sep=",", header=TRUE)
df4_cd3cd28g0076 <- read.table(file="4. cd3cd28g0076.csv",sep=",", header=TRUE)
df5_cd3cd28psitect <- read.table(file="5. cd3cd28psitect.csv",sep=",", header=TRUE)
df6_cd3cd28u0126 <- read.table(file="6. cd3cd28u0126.csv",sep=",", header=TRUE)
df7_cd3cd28ly <- read.table(file="7. cd3cd28ly.csv",sep=",", header=TRUE)
df8_pma <- read.table(file="8. pma.csv",sep=",", header=TRUE)
df9_b2camp <- read.table(file="9. b2camp.csv",sep=",", header=TRUE)
df10_cd3cd28icam2aktinhib <- read.table(file="10. cd3cd28icam2aktinhib.csv",sep=",", header=TRUE)
df11_cd3cd28icam2g0076 <- read.table(file="11. cd3cd28icam2g0076.csv",sep=",", header=TRUE)
df12_cd3cd28icam2psit <- read.table(file="12. cd3cd28icam2psit.csv",sep=",", header=TRUE)
df13_cd3cd28icam2u0126 <- read.table(file="13. cd3cd28icam2u0126.csv",sep=",", header=TRUE)
df14_cd3cd28icam2ly <- read.table(file="14. cd3cd28icam2ly.csv",sep=",", header=TRUE)
```

## Column names

Only df8 contained two columns with different names (in lower case). 
I unfied column names.

```{r message=FALSE}
df8_pma <- df8_pma %>% rename (PIP2 = pip2, PIP3 = pip3)
```

## New column for perturbations
Measurments are obtained from two different perturbations ("general perturbation" : GP1 and GP2). At the same time, these perturbations are combined with different treatments (considered below).
I added a new column called GP to each table to classify the data depending on the general perturbation is applied : GP  = 1 for GP1 and GP = 2 for GP2.

```{r message=FALSE}
df1_cd3cd28 <- df1_cd3cd28 %>% 
              mutate(treatment = "cd3cd28") %>% mutate (GP = 1) 
df2_cd3cd28icam2 <- df2_cd3cd28icam2 %>% 
                    mutate(treatment = "cd3cd28icam2") %>% mutate (GP = 2)
df3_cd3cd28aktinhib <- df3_cd3cd28aktinhib %>% 
                      mutate(treatment = "cd3cd28aktinhib") %>% mutate (GP = 1) 
df4_cd3cd28g0076 <- df4_cd3cd28g0076 %>% 
                    mutate(treatment = "cd3cd28g0076") %>% mutate (GP = 1) 
df5_cd3cd28psitect <- df5_cd3cd28psitect %>% 
                      mutate(treatment = "cd3cd28psitect") %>% mutate (GP = 1) 
df6_cd3cd28u0126 <- df6_cd3cd28u0126 %>% 
                    mutate(treatment = "cd3cd28u0126") %>% mutate (GP = 1) 
df7_cd3cd28ly <- df7_cd3cd28ly %>% 
                  mutate(treatment = "cd3cd28ly") %>% mutate (GP = 1) 
df8_pma <- df8_pma %>% 
            mutate(treatment = "pma") %>% mutate (GP = 1) 
df9_b2camp <- df9_b2camp %>% 
              mutate(treatment = "b2camp") %>% mutate (GP = 1) 
df10_cd3cd28icam2aktinhib <-df10_cd3cd28icam2aktinhib %>% 
                            mutate(treatment = "cd3cd28icam2aktinhib") %>% mutate (GP = 2)
df11_cd3cd28icam2g0076 <-df11_cd3cd28icam2g0076 %>% 
                          mutate(treatment = "cd3cd28icam2g0076") %>% mutate (GP = 2)
df12_cd3cd28icam2psit <- df12_cd3cd28icam2psit %>% 
                        mutate(treatment = "cd3cd28icam2psit") %>% mutate (GP = 2)
df13_cd3cd28icam2u0126 <- df13_cd3cd28icam2u0126 %>% 
                          mutate(treatment = "cd3cd28icam2u0126") %>% mutate (GP = 2)
df14_cd3cd28icam2ly <- df14_cd3cd28icam2ly %>% 
                        mutate(treatment = "cd3cd28icam2ly") %>% mutate (GP = 2)
```

## Unique table
I created a unique table for all the perturbations/treatments.

```{r message=FALSE}
alldf <- bind_rows(df1_cd3cd28, df2_cd3cd28icam2, df3_cd3cd28aktinhib, df4_cd3cd28g0076,      df5_cd3cd28psitect, df6_cd3cd28u0126, df7_cd3cd28ly, df8_pma, df9_b2camp, df10_cd3cd28icam2aktinhib, df11_cd3cd28icam2g0076, df12_cd3cd28icam2psit, df13_cd3cd28icam2u0126, df14_cd3cd28icam2ly)
```

## Dummy variables for the different treatments
I created dummy variables for each of the treatments (independently of the perturbation)

```{r message=FALSE}
alldf <- alldf %>%
          mutate (Akt_inh1 = ifelse (treatment == c("cd3cd28aktinhib", "cd3cd28icam2aktinhib"), 1, 0)) %>%
          mutate (PKC_inh = ifelse (treatment == c("cd3cd28g0076", "cd3cd28icam2g0076"), 1, 0)) %>%
          mutate (PIP2_inh = ifelse (treatment == c("cd3cd28psitect", "cd3cd28icam2psit"), 1, 0)) %>%
          mutate (MEK_inh = ifelse (treatment == c("cd3cd28u0126", "cd3cd28icam2u0126"), 1, 0)) %>%
          mutate (Akt_inh2 = ifelse (treatment == c("cd3cd28ly", "cd3cd28icam2ly"), 1, 0)) %>%
          mutate (PKC_act = ifelse (treatment == "pma", 1, 0)) %>%
          mutate (PKA_act = ifelse (treatment == "b2camp", 1, 0)) 
```

## Reorder columns
I reorder the columns to have the treatment names and dummy variables first, and after that all the measurments done.

```{r message=FALSE}
alldf <- alldf %>% select(treatment, GP, Akt_inh1, PKC_inh, PIP2_inh, MEK_inh, Akt_inh2, PKC_act, PKA_act, everything())
```

## Save table
I saved the tabble as .csv document (capstone_project.csv)

```{r message=FALSE}
write.table(alldf, file = "capstone_project.csv", sep = ",", col.names = NA)

```


